\textbf{Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata}

### 2023-05-08 manuscript_v5/02-methods.qmd
new directory for version, tidy code

Preliminary analyses of the data included the removal of outliers, of observations where respondents did not provide answers in all the covariates of interest, matching the city of each observations with its corresponding LHIN and Health Region, and removing observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions). After all the preliminary analyses indicated above, the total number of observations used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021. The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.

@@

Comparisons between the variables of the survey were performed using the $\chi^2$ test. We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data. The reference groups in the model were: 16-34 years (age group), October(Month), CAD 60,000 and above (income), White/Caucasian (Race), Toronto (Health Region). Details regarding the correction can be found in the Appendix. All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], `quarto` [@quarto], `modelsummary` [@modelsummary], and `gtsummary` [@gtsummary].

>>>>

Preliminary analyses of the data included the removal of outliers (**should we still do this? it's only 19 observations with income >110k and household of 1, but we are not even using such income bracket in the analysis because we re-grouped the data, and the household size variable has 90% missing rate**), of observations where respondents did not provide answers in all the covariates of interest, matching the city of each observations with its corresponding LHIN and Health Region, and removing observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions). After all the preliminary analyses indicated above, the total number of observations used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021. The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.

@@

We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data. Details regarding the correction can be found in the Appendix. All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], `quarto` [@quarto], `modelsummary` [@modelsummary], and `gtsummary` [@gtsummary].

### 2023-05-08 manuscript_v5/03-results.qmd
new directory for version, tidy code

@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status. A total of 3,549 observations were used for analysis, from which 27\% (958) corresponded to individuals that reported not having received the first dose of the vaccine. Respondents between 16 and 34 years of age, and those that lived within the Health Region of Toronto had the highest number of observations in the data (42.8\% and 37.3\% of the total, respectively). Moreover, individuals who identified as White or Caucasian reported the highest vaccination uptake (82\%), in contrast to respondents who identified as which had had the lowest vaccination uptake rate (60\%) across all groups. From all the covariates, significant associations were identified between income, race, and vaccination rate.

@@

source(here::here("code","data_preparation.R"))

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% add_p(test = everything () ~ "chisq.test")%>%
  bold_labels() %>%
  modify_table_styling(
    columns=label,
    rows=label=="Other",
    footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
  )

# data_summary %>%
#   modify_header(label = "**Variable**") %>%
#   as_kable_extra(booktabs = TRUE,linesep="") %>%
#   # reduce font size to make table fit. 
#   # you may also use the `latex_options = "scale_down"` argument here.
#   kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)

data_summary %>%
  modify_header(label = "**Variable**") %>%
  as_gt()

# test<-apply(xtabs(~first_dose_m + Race,clean_data), 2, chisq.test)
# 
# 
# clean_data %>%
#   filter(income_ord %in% c("60000 and above"))%>%
#   tbl_summary(percent="row",
#               by= first_dose_m,
#               missing="no",
#               include=Health_Region,
#               label=list(
#                  income_ord="Income (CAD)")
#   ) %>% add_p()%>%
#   bold_labels()

# clean_data %>%
#   filter(income_ord %in% c("25000-59999"))%>%
#   tbl_summary(percent="row",
#               by= income_ord,
#               missing="no",
#               include=(income_ord),
#               label=list(
#                  income_ord="Income (CAD)")
#   ) %>% add_p()%>%
#   bold_labels()

@@

#call the script for raking
source(here::here("code","raking.R"))

#logistic regression
source(here::here("code","regression.R"))

#create dataframe from modelplot, remove names of factors, drop the intercept

mplot1<-modelplot(m0,draw=FALSE,
                  coef_omit = "Intercept", exponentiate = TRUE) %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

#add column by factor 
mplot1$factor<-c(rep("Age",2),
                 rep("Month",2),
                 rep("Income",2),
                 rep("Race/Ethnicity",8),
                 rep("Health Region",3),
                 rep("Income and\nRace/Ethnicity",16),
                 rep("Race/Ethnicity\nand\nHealth Region",24)
                 )

#order factors for plot
#mplot1$term_ord <- factor(mplot1$term,ordered=TRUE,levels=mplot1$term)

mplot1$factor <- factor(mplot1$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))

# remove non-significant interaction factors

test<-mplot1%>%
  filter(str_detect(term,':') & p.value>0.05)

#remove non-significant interactions from dataframe
mplot1<-anti_join(mplot1,test)

mplot1<-mplot1%>%
mutate(term=str_replace_all(term,':'," and "))

#make plot
mplot1 %>% 
ggplot(aes(y=term))+
  geom_vline(xintercept = 1, linetype="dashed",color="red") +
  geom_point(aes(x=estimate),size=2)+
  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
  scale_y_discrete(limits=rev)+
  xlab("")+
  ylab("")+
  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y",labeller = label_wrap_gen(multi_line = TRUE))+
  theme_classic()
```

@@

#create dataframe from modelplot, remove names of factors, drop the intercept

mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))

#add column by factor 
mplot2$factor<-c(rep("Age",2),
                 rep("Month",2),
                 rep("Income",2),
                 rep("Race/Ethnicity",8),
                 rep("Health Region",3),
                 rep("Income and\nRace/Ethnicity",16),
                 rep("Race/Ethnicity\nand\nHealth Region",24)
                 )

#order factors for plot
#mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)

mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))

# remove non-significant interaction factors

test<-mplot2%>%
  filter(str_detect(term,':') & p.value>0.05)

mplot2<-anti_join(mplot2,test)

mplot2<-mplot2%>%
mutate(term=str_replace_all(term,':'," and "))

mplot2 %>% 
ggplot(aes(y=term))+
  geom_vline(xintercept = 1, linetype="dashed",color="red") +
  geom_point(aes(x=estimate),size=2)+
  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
  scale_y_discrete(limits=rev)+
  xlab("")+
  ylab("")+
  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
  theme_classic()
```

>>>>

@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status. A total of **6,236** observations were used for analysis, from which 24\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine. Across the socio-demographic characteristics collected by the survey, higher vaccination uptake was reported by individuals in the highest income bracket considered, those between 16 and 34 years of age, individuals that lived in the East Health Region, and individuals that identified as White or Caucasian. During the months covered by the survey, the highest vaccination uptake was observed in January of 2022.

@@

source(here::here("code","data_preparation_May_08.R"))

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% add_p(test = everything () ~ "chisq.test")%>%
  bold_labels() %>%
  modify_table_styling(
    columns=label,
    rows=label=="Other",
    footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
  ) %>%
  modify_header(label = "**Variable**") %>%
  as_gt()

data_summary

@@

#call the script for raking
source(here("code","raking.R"))

#logistic regression
source(here("code","regression.R"))

# call script that creates figure for model

source(here("code","uncorr_model.R"))

#call figure created in the script above
uncorr_mod

@@

#call script with code for figure
source(here("code","corr_model.R"))

#call figure
corr_mod
```
